Predictive modeling using genomic annotations

### 2018-11-14 Manuscript/manuscript.Rmd
updated manuscript

In this study we have developed an ensemble framework that aims to address and compare a variety of data irregularities and characteristics that are associated with the prediction of TAD boundaries. We evaluated domain data obtained from four different resolution contact matrices for the GM12878 cell line including 5kb, 25kb, 50kb, and 100kb. For each resolution, we examined the inclusion of three different predictor types as the feature space for downstream modelling. We coupled each predictor type with two separate variable reduction techniques. Lastly, we performed a series of three different re-sampling techniques aimed at created balanced classes. Random forest classification algorithms were performed to both compare each combination of the above ensemble, and investigate which functional genomic features were most associated with the formation of TAD boundaries.  Model performance was evaluated based on a total of four metrics which included test accuracy, area under the ROC curve (AUC), F1 score, and Matthew's Correlation Coefficient (MCC).

@@

>>>>

In this study we have developed an ensemble framework that aims to address and compare a variety of data irregularities and characteristics that are associated with the prediction of TAD boundaries. We evaluated domain data obtained from four different resolution contact matrices for the GM12878 cell line including 5kb, 25kb, 50kb, and 100kb. For each resolution, we examined the inclusion of three different predictor types as the feature space for downstream modelling. We coupled each predictor type with two separate variable reduction techniques. Lastly, we performed a series of three different re-sampling techniques aimed at created balanced classes. Random forest classification algorithms were performed to both compare each combination of the above ensemble, and investigate which functional genomic features were most associated with the formation of TAD boundaries.  Model performance was evaluated based on a total of four metrics which included test accuracy, area under the ROC curve (AUC), F1 score, and Matthew's Correlation Coefficient (MCC). [mention reasons for including 4 metrics]

@@

###DISTANCE PREDICTORS IMPROVE MODEL PERFORMANCE IN ALL RESOLUTIONS

Figures 4-7 present the results across all of the ensemble systems for each resolution of domain data. Focusing on the top row in each of Figure 4-7, indicating no balancing, we see that the model using distance predictors, represented by the blue bars, outperformed models using the other two predictor types. However, the disparity become less evident as the resolution increased. In each of the 4 figures we see that the F1-score and MCC metrics were noticeably lower than the accuracy and AUC. This highlights the problems associated with applying predictive models to unbalanced data.

###CLASS BALANCING IMPROVES MODEL PERFORMANCE IN ALL RESOLUTIONS

Moving down the rows in each of Figures 4-7 we see that model performance was markedly improved among the F1-scores and MCCs after creating balanced classes compared to the results using no re-sampling. For each of the resolutions, we see that random under-sampling and SMOTE re-sampling techniques performed very similarly, while random over-sampling performed poorly in comparison. Furthermore, we see that the elastic-net reduction technique outperformed the LASSO. Although, the disparity became less evident as the resolution increased.

###PERFORMANCES BECOME INVARIANT AS RESOLUTION INCREASES
